---
title: "Homework3"
author: "Andrea Domen"
date: "`r Sys.Date()`"
output: pdf_document
knit: (function(inputFile, encoding) {
      out_dir <- "../Outputs";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Sleuth3)
```

# Q1

**Ch 19 Ex 11** <br>

```{r}
crab <- tibble(crab_sp = c("Red crab", "Dungeness crab", "marginals"),
               parasite = c(5, 0, 5),
               no_parasite = c(312, 503, 815),
               marginals = c(317, 503, 820))
```

## a

```{r}
red.parasite.exp <- (crab$marginals[1]*crab$parasite[3])/crab$marginals[3]
paste("The expected value for red crabs with parasites is", round(red.parasite.exp, 4))
```

$H_0: \frac{\omega_{red\; crab\; parasite}}{\omega_{dungeness\; crab\; parasite}} =1$ <br>

$H_a: \frac{\omega_{red\; crab\; parasite}}{\omega_{dungeness\; crab\; parasite}} >1$

Since $Observed > Expected$, but the minimum of the marginals is 5, only $P(5)$ needs to be calculated.

## b

```{r}
# great, the table I made above won't work for fisher.test so I guess I get to make it again
crab.matrix <- matrix(c(5, 0, 312, 503), 2, 2)
fisher.test(crab.matrix, alternative = "greater")
```

## c

There is strong evidence to suggest that Red crabs have significantly higher odds of being infected with parasites ($p = 0.008$) with 5 out of 317 (1.6%) of Red crabs infected compared with 0 out of 503 Dungeness crabs. As there were no parasites on the Dungeness crabs, an odds ratio cannot be reported.

\newpage

# Q2

**Ch 19 Ex 12** <br>

```{r}
hunt <- tibble(behavior = c("Hunt alone", "Group hunt"),
               single_prey = c(17, 14),
               mult_prey = c(12, 7))
```

## a

```{r}
hunt.matrix <- matrix(c(17, 14, 12, 7), 2, 2)
prop.test(hunt.matrix, alternative = "two.sided")
```

$H_0: hunting\; strategies\; and\; prey\; are\; independent$ <br>

$H_a: hunting\; strategies\; and\; prey\; are\; not\; independent$

The hunting strategy (alone or in groups) is and prey type are independent from one another, with no significant association between the two ($p=0.78$).

## b

$H_0: \frac{\omega_{hunt\; strategy}}{\omega_{prey\; sought}} =1$ <br>

$H_a: \frac{\omega_{hunt\; strategy}}{\omega_{prey\; sought}} \neq 1$

```{r}
fisher.test(hunt.matrix, alternative = "two.sided")
```

Using Fisher's exact test, the hunting strategy (alone or in groups) is and prey type are independent from one another, with no significant association between the two ($p=0.77$). 

\newpage

# Q3

**Ch 19 Ex 18** <br>

$H_0: \frac{\omega_{i,\; female\; survival}}{\omega_{i,\; male\; survival}} = 1$ when controlling for age. <br>

$H_a: \frac{\omega_{i,\; female\; survival}}{\omega_{i,\; male\; survival}} > 1$ when controlling for age.

```{r}
donner <- ex1918

donner.array <- array(NA, c(2, 2, 6))

for (k in 1:6){
  donner.array[, , k] <- as.matrix(donner[(2*(k-1) +1) : (2*k), 3:4])
}

mantelhaen.test(donner.array)
```

With a common odds ratio of 0.258, there is no evidence that the odds of survival were higher for females than males when controlling for age ($p=0.09$). However, this is a borderline value at the chosen significance level.
